Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2021-01-30 paper/paper.tex
paper: Review Abstract

\resumen{
The characterization and understanding of \emph{Open Clusters}
allow us to better understand properties and mechanisms about the Universe
such as stellar formation and the kind of regions where these events occur.
They also provide information about stellar processes and the evolution of the galactic disk.

In this paper, we develop a method to characterize open clusters by using
\emph{artificial intelligence} tools, such as clustering by \emph{K-Means}
and clustering based on \emph{Artificial Neural Networks} by implementing
the \emph{Deep Embedded Clustering} model.
We are using \emph{Gaia DR2 database} as data source for testing our models.

The developed method aims to improve current techniques of the state of the art.
We achieve improvements not only in terms of \emph{computational efficiency},
with lower computational requirements, but in \emph{usability},
reducing the number of hyperparameters to obtain a good characterization of the analyzed clusters.

>>>>

\resumen{
The characterization and understanding of \emph{Open Clusters}
allow us to better understand properties and mechanisms about the Universe
such as stellar formation and the regions where these events occur.
They also provide information about stellar processes and the evolution of the galactic disk.

In this paper, we develop a method to characterize open clusters by using
\emph{machine learning} tools, such as clustering by \emph{K-Means}
and clustering based on \emph{Artificial Neural Networks} by adapting
the \emph{Deep Embedded Clustering} model.
We are using \emph{Gaia DR2 database} as data source.

The developed method aims to improve current techniques of the state of the art.
We achieve improvements not only in terms of \emph{computational efficiency},
with lower computational requirements, but in \emph{usability}, reducing the number
of hyperparameters to get a good characterization of the analyzed clusters.
